---
title: "Leaflet"
author: "Kushal Kharel"
date: "8/27/2022"
output: html_document
---

Leaftlet is JS library, have R package for it. Can create interactive maps (creating JS widget actually)

## My first map

```{r, eval = FALSE}

library(leaflet) 
my_map = leaflet() %>% #leaflet function creates a map widget& we store it in a variable so we can modify it later
  addTiles() # add tiles function adds mapping data from open street map
my_map

```

Adding markers to the map one at a time by specifying longitude and latitude.

Can specify popup text when you click on marker with popup argument

```{r}
library(leaflet)

my_map = leaflet() %>%
  addTiles() %>%
  addMarkers(lat = 39.2980803, lng = -76.5898801,
             popup = "Johns Hopkins Bloomberg School of Public Health (Oldest School in the country")
my_map

```

Adding marker one at a time is not practical. If we have df with lat and long column then we can pipe the 
df into leaflet() to add all the points at once.

```{r}
set.seed(2016-04-25)

df = data.frame(lat = runif(20, min = 39.2, max = 39.3),
                lng = runif(20, min = -76.6, max = -76.5))

df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```


We can make our own markers from .png files

```{r, echo = FALSE}
# making custom markers 
library(leaflet)
hopkinsIcon = makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, 
  iconHeight = 31,
  iconAnchorX = 31*215/230/2,
  iconAnchorY = 16
)

hopkinsLatLong = data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929768, -76.6206598, -76.5469683, -76.6150537, -76.6016766)
)

hopkinsLatLong %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()#(icon = hopkinsIcon)
```

We may want to add popup to each marker. We can either provide plain text or html to be rendered inside of each popup

```{r, eval = FALSE}

hopkinsIcon = makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, 
  iconHeight = 31,
  iconAnchorX = 31*215/230/2,
  iconAnchorY = 16
)

hopkinsLatLong = data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929768, -76.6206598, -76.5469683, -76.6150537, -76.6016766)
)


hopkinsSites = c(
  "<a href='http://www.jhsph.edu/'>East Balitmore Campus</a",
  "<a href = 'https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>"
)

hopkinsLatLong %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = hopkinsSites)#(icon = hopkinsIcon)


```


## Mapping Clusters

Sometimes we may have so many points on a map. It doesn't make sense to plot every marker.
We can instead plot clusters of markers using clusterOpetions = markerClusterOptions().
When we zoom in t oeach cluster, the clusters will separate until we see individual markers

```{r}
df = data.frame(lat = runif(500, min = 39.25, max = 39.35),
                lng = runif(500, min = -76.65, max = -76.55))

df %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

## Mapping Circle Markers

Instead of adding markers or clusters, we can add circle markers using `addCircleMarkers()`

```{r, eval = FALSE}
df = data.frame(lat = runif(20, min = 39.2, max = 39.3),
                lng = runif(20, min = -76.6, max = -76.5))

df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()

```

## Drawing Circles

We can also draw arbitrary shapes on the maps we create.
Lets draw a map where circle on each city is proportional to the population of that city

```{r}
md_cities = data.frame(names = c("Baltimore", "Frederick", "Rockville"),
                       pop = c(63433, 66168, 62332),
                       lat = c(39.2920592, 39.4143921, 39.0840),
                       lng = c(-76.6077852, -77.4204875, -77.1528))

md_cities %>%
  leaflet() %>%
  addTiles() %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop)*30)
```

```{r}
df = data.frame(latitude = runif(20, min = 39.2, max = 39.3),
                longitude = runif(20, min = -76.6, max = -76.5),
                   col = sample(x = c("red", "green", "yellow"), 20, replace = TRUE),
                   stringsAsFactors = FALSE)

df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("red", "green", "yellow"))
```


```{r}
df = data.frame(lat = runif(20, min = 39.2, max = 39.3),
                lng = runif(20, min = -76.6, max = -76.5))

df %>%
  leaflet() %>%
  addTiles() %>%
  addRectangles( lat1 =  df$lat[1],
                 lat2 = df$lat[2],
                 lng1 = df$lng[1],
                 lng2 =  df$lng[2])
```

